\documentclass[a4paper, 12pt, titlepage]{article}

% Including needed packages
\usepackage[margin=2cm]{geometry}
\usepackage{amsmath}

\title
{{\em INF 551 - Computer-aided reasoning}\\
Mini-project - Encoding funny games into SAT\\
{\bf Report}}
\author{ROHRMANN Till \and GROSSHANS Nathan}
\date{}

\begin{document}

\maketitle


\section{Introduction}
Nowadays, computers became a powerful tool to help us humans in a variety of ways.
We humans only have to describe the problem in such a way that the computer can understand it and process it.
Instead of developing for every issue a proprietary solution it is often more efficient to use existing tools which are well understood and efficiently implemented.
One of these general problems is SAT for which plenty of solvers have already been developed.
Thus, one only has to find a reduction from the original problem to a Boolean formula which one then tries to solve.
In the context of this work we investigate the applicability of such a reduction for the two games Skyscraper and Hex-a-Hop and evaluate the results.

In the first section we present the game Skyscraper and describe its reduction to a SAT problem.
In the second section we do the same for the game Hex-a-Hop.

\section{Skyscraper}
The skyscraper game is a little logical game that works like the Sudoku game.
Having a square grid of a fixed size $n$, the goal is to place skyscrapers of
different heights (from 1 to the size of the grid) in the squares of the grid,
such that a skyscraper of each height appears once and only once on each row and
each column, and that the visibility constraints are respected. The visibility
constraints are given by numbers, one on each side of a row or column,
indicating how many buildings must be seen when looking from the given side
towards the skyscraper row or column. There can also be no constraint at all for
a given row or column side.

\subsection{The encoding}
In order to encode the game rules and to obtain the final result we introduced Boolean variables $v_{x,y}^h$ saying whether a skyscraper of the size $h$ is on the field $(x,y)$ or not. The game rules are then realized by the following formulas:
	\begin{itemize}
		\item There has to be at least one skyscraper on each field
			\begin{eqnarray}
				\pmb{atLeastOne}&\equiv&\bigwedge_{x=0,y=0}^{n-1} \left(\bigvee_{h=0}^{n-1} v_{x,y}^h \right)
			\end{eqnarray}
		\item There can be at most one skyscraper on each field
			\begin{eqnarray}
				\pmb{atMostOne}&\equiv&\bigwedge_{x=0,y=0}^{n-1} \left( \bigwedge_{h_\not=h_2}^{n-2}\neg(v_{x,y}^{h_1} \wedge v_{x,y}^{h_2})  \right)
			\end{eqnarray}
		\item There can be only one skyscraper of a certain height in each row and column
			\begin{eqnarray}
				\pmb{oneInARow} &\equiv& \bigwedge_{x=0}^{n-1} \left( \bigwedge_{h=0}^{n-1} \left( \bigwedge_{y_1 \not = y_2}^{n-1} \neg(v_{x,y_1}^h \wedge v_{x,y_2}^h) \right) \right) \\
				\pmb{oneInAColumn} &\equiv&  \bigwedge_{y=0}^{n-1} \left( \bigwedge_{h=0}^{n-1} \left( \bigwedge_{x_1 \not = x_2}^{n-1} \neg(v_{x_1,y}^h \wedge v_{x_2,y}^h) \right) \right)
			\end{eqnarray}
	\end{itemize}

In order to realize the constraints, we introduced two predicates $maxSkyscraper$ and $numberSkyscraper$. The first predicate saves which is the highest skyscraper seen so far and the latter counts the number of seen skyscraper up to a given field. Let $D=\{N, E, W, S\}$ be the set of directions. $maxSkyscraper_{d}(x,y,h)$ with $d \in D$ is true iff seen from the field $(x,y)$ in the direction $d$, excluding itself, the highest skyscraper has the height $h-1$. $numberSkyscraper_{d}(x,y,n)$ with $d \in D$ is true iff one can see $n$ skyscrapers from a side in the direction of $d$ up to the field $(x,y)$. For the purpose of readability, we are going to abbreviate $maxSkyscraper$ by $mS$ and $numberSkyscraper$ by $nS$. Regarding the predicates as functions, then they are defined by:
	\begin{eqnarray}
		mS_{d}(\pmb x) &=& 
		\begin{cases}
			height(\pmb x)+1 & height(\pmb x) \ge mS_{d}(\pmb x + \pmb e_{d})\\
			mS_{d}(\pmb x + \pmb e_{d}) &\text{if not}
		\end{cases}
		\\
		nS_{d}(\pmb x) &=&
		\begin{cases}
			nS_{d}(\pmb x - \pmb e_{d}) + 1 & height(\pmb x) \ge mS_{d}(\pmb x)\\
			nS_{d}(\pmb x - \pmb e_{d}) &\text{sinon}
		\end{cases}
	\end{eqnarray}
with $\pmb e_{N} = (-1,0)^T, \pmb e_{E} = (0,1)^T, \pmb e_{S} = (1,0)^T,\pmb e_{W} = (0,-1)^T$ and $\pmb x = (x,y)^T$ being the position vector. Moreover, we set the $mS_{d}$ predicate on the respective side to $0$ and the $nS_{d}$ predicate to $1$. The behavior of $mS$ is modeled by the following formulas:
\begin{itemize}
	\item $mS$ can take only at most one value
	\begin{eqnarray}
		\pmb{mSisAFunction} &\equiv& \bigwedge_{d\in D} \left( \bigwedge_{x=0,y=0}^{n-1} \left( \bigwedge_{v_1 \not = v_2} \neg \left( mS_{d}(x,y,v_1) \wedge mS_{d}(x,y,v_2) \right) \right) \right)
	\end{eqnarray}
	\item Set $mS$ on the borders to zero
	\begin{eqnarray}
		\pmb{mStoZero} &\equiv& \bigwedge_{x=0}^{n-1} mS_{N}(0,x,0) \wedge mS_{E}(x,n-1,0) \wedge \nonumber \\
	&&  mS_{S}(n-1,x,0) \wedge mS_{W}(x,0,0)
	\end{eqnarray}
	\item Value propagation if the skyscraper on the field is smaller than $mS$.
	\begin{eqnarray}
		\pmb{mSvPropC} &\equiv& \bigwedge_{d \in D,x=0,y=0} \bigwedge_{h_1=0}^{n-1} \bigwedge_{h_2 = h_1+1}^{n} \left( v_{x,y}^{h_1} \wedge mS_{d}(\pmb x + \pmb e_{d},h_2) \Rightarrow mS_{d}(\pmb x,h_2) \right)
	\end{eqnarray}
	\item Value propagation if the skyscraper on the field is greater or equal than $mS$.
	\begin{eqnarray}
		\pmb{mSvPropAug} &\equiv& \bigwedge_{d \in D,x=0,y=0} \bigwedge_{h_1=0}^{n-1} \bigwedge_{h_2=h_1}^{n-1} \left( v_{x,y}^{h_2} \wedge mS_{d}(\pmb x + \pmb e_{d},h_1) \Rightarrow mS_{d}(\pmb x,h_2+1) \right)
	\end{eqnarray}
\end{itemize}

The behavior of $nS$ is given by:
\begin{itemize}
	
\item $nS$ can take only at most one value
	\begin{eqnarray}
		\pmb{nSisAFunction} &\equiv& \bigwedge_{d\in D} \left( \bigwedge_{x=0,y=0}^{n-1} \left( \bigwedge_{v_1 \not = v_2} \neg \left( nS_{d}(x,y,v_1) \wedge nS_{d}(x,y,v_2) \right) \right) \right)
	\end{eqnarray}
	
\item Set $nS$ on the border to one
	\begin{eqnarray}
		\pmb{nStoOne} &\equiv& \bigwedge_{x=0}^{n-1} nS_{N}(n-1,x,1) \wedge nS_{E}(x,0,1) \wedge \nonumber \\
	&&  nS_{S}(0,x,1) \wedge nS_{W}(x,n-1,1)
	\end{eqnarray}
	
\item Value propagation if the skyscraper on the field is smaller than $mS$
	\begin{eqnarray}
		\pmb{nSvPropC} &\equiv& \bigwedge_{\substack{d\in D\\x=0\\y=0}} \bigwedge_{h_{1}=0}^{n-1} \bigwedge_{\substack{h_{2} = h_{1}+1\\h_{3}=0}}^{n} \left( v_{x,y}^{h_1} \wedge mS_d(\pmb x,h_2) \wedge nS_{d^{-1}}(\pmb x - \pmb e_d,h_3) \right. \nonumber \\
	&& \left.\Rightarrow nS_{d^{-1}}(\pmb x - \pmb e_d,h_3) \right)
	\end{eqnarray}
	with $d^{-1}$ being the inverse direction.

	\item Value propagation if the skyscraper is greater or equal than $mS$
	\begin{eqnarray}
		\pmb{nSvPropAug} &\equiv& \bigwedge_{\substack{d\in D\\x=0\\y=0}} \bigwedge_{h_{1}=0}^{n-1} \bigwedge_{\substack{h_{2} = h_{1}\\h_{3}=0}}^{n} \left( v_{x,y}^{h_2} \wedge mS_d(\pmb x,h_1) \wedge nS_{d^{-1}}(\pmb x - \pmb e_d,h_3) \right. \nonumber \\
	&& \left.\Rightarrow nS_{d^{-1}}(\pmb x - \pmb e_d,h_3+1) \right)
	\end{eqnarray}
\end{itemize}

Since all of these formulas are already in CNF form there is no transformation needed. In the last step, the constraints will be specified. This can be easily achieved by constraining the $nS$ at the respective positions to be true for the constraint values. For example, assuming that we have the constraint $c$ for the first line on the left side, then it would be realized by adding $nS_{E}(0,n-1,c)$ as a new clause.

It is noteworthy that by introducing the two additional predicates, we have achieved a polynomial SAT reduction in the size of the field side length $n$. In order to encode the field with the different skyscrapers we need $n^3$ variables. For the two predicates another $2\cdot n^2\cdot(n+1)$ variables are required. Thus the complexity of our encoding is in $O(n^3)$.

\subsection{Evaluation}

With our encoding we could solve all levels within a fraction of a second. Even the levels, where not all constraints were specified, did not pose a problem to our encoding. The produced results were verified by inserting them into the game. However, one has to admit that due to the lack of dynamic the encoding is rather simple and does not require many variables. Thus, it is easy to solve for up to date SAT solvers.

\subsection{How to use the program}
The encoding is realized by the python program encoding.py. The only prerequisite to run the program is an installed python version. The program awaits an input file which specifies the field and the constraints. The first line contains the side length $n$ of the field. It is followed by $2\cdot n$ pairs of numbers, separated by a white space, specifying the constraints. The range for the constraints is $[0,\ldots,n]$ whereby $0$ denotes that there is no constraint. The first $n$ pairs specify the constraints for the lines, the first number indicating the left and the second number indicating the right constraint. The last $n$ pairs specify the column constraints, the first number indicating the top and the second number indicating the bottom constraint. The program outputs the encoding in the DIMACS format to the standard output and can thus be easily given to a third-party SAT solver capable of reading the DIMACS format.

We used the Glucose SAT solver for our project because it could solve sufficiently fast our problems. But any other SAT solver can be used as well.

The output of the SAT solver can be visualized by the python program decoding.py. The program expects as input arguments the file defining the field and the model calculated by the SAT solver. Given that, it will print the placement of the skyscraper on the field for the given constraints.

For the sake of simplicity, we recommend to use the script \emph{skyscraper.sh} which takes as single argument the field file and combines all of the aforementioned steps. The only thing one has to adapt is the path to the SAT solver.



\section{Hex-a-Hop}
Hex-a-Hop is a puzzle-based game with hundred levels where the player jumps from
one hexagonal tile to another. The goal, for each level, is to destroy all the
green tiles by visiting them once, without getting trapped or falling into the
water. The levels gradually become more and more difficult, unveiling more and
more new tile types having different behaviors (trampolines, high tiles...).

\subsection{The encoding}

\subsubsection{Representing a level}
A level can be represented by a grid (field) of tiles of different types. As
these tiles are hexagonal, the representation is a bit more complicated as if
they where just squares. In fact, as two successive rows of tiles are
"interleaved" in the sense that the tiles in a row appear each two columns
alternately with the tiles in the adjacent rows, we can represent two rows of
the level in one grid line where the tiles of one row are in the columns with
odd column indexes and the tiles of the other row in the columns with even
column indexes. Naturally, we must then be careful when doing movements around
the grid (the function {\em getDirectionVector} therefore gives the right
movement to do on the grid when moving in a given direction).


\subsubsection{The dynamic type of tiles}
In our first version of the encoding, there were no more variables than those
encoding the position at each time step. But it is convenient to add some other
variables, auxiliary variables, that help to memorize some useful states.
Dynamic type variables therefore memorize the type a tile has at a given
time step, as it can be green, turquoise, destroyed... They are of the form
$(x, y, t, dynamic\_type)$ where in our case $dynamic\_type$ can take the values
$d_{green}$, $d_{turquoise}$ or $d_{destroyed}$. These variables must be
initialized correctly thanks to the initial state in the level, and their value
must be correctly updated. Therefore, for each position and time step
$(x, y, t)$, we have the following implications (clauses) that cover all the
possible cases :
\begin{equation}
\neg(x, y, t-1, d_{turquoise}) \Rightarrow (x, y, t, d_{turquoise})
\end{equation}
\begin{equation}
(x, y, t-1, d_{turquoise}) \wedge (x, y, t-1) \Rightarrow
					      \neg(x, y, t, d_{turquoise})
\end{equation}
\begin{equation}
(x, y, t-1, d_{turquoise}) \wedge \neg(x, y, t-1) \Rightarrow
						  (x, y, t, d_{turquoise})
\end{equation}
\begin{equation}
\neg(x, y, t-1, d_{green}) \wedge \neg(x, y, t-1, d_{turquoise}) \Rightarrow
							\neg(x, y, t, d_{green})
\end{equation}
\begin{equation}
(x, y, t-1, d_{turquoise}) \wedge (x, y, t-1) \Rightarrow (x, y, t, d_{green})
\end{equation}
\begin{equation}
(x, y, t-1, d_{turquoise}) \wedge \neg(x, y, t-1) \Rightarrow
						  \neg(x, y, t, d_{green})
\end{equation}
\begin{equation}
(x, y, t-1, d_{green}) \wedge (x, y, t-1) \Rightarrow \neg(x, y, t, d_{green})
\end{equation}
\begin{equation}
(x, y, t-1, d_{green}) \wedge \neg(x, y, t-1) \Rightarrow (x, y, t, d_{green})
\end{equation}
\begin{equation}
(x, y, t-1, d_{destroyed}) \Rightarrow (x, y, t, d_{destroyed})
\end{equation}
\begin{equation}
(x, y, t-1, d_{green}) \wedge \neg(x, y, t-1) \Rightarrow
					      (x, y, t, d_{destroyed})
\end{equation}


\subsubsection{The main idea of the encoding}
The problem of deciding whether a Hex-a-Hop level can be solved can be
transformed into the parameterized problem of deciding whether a Hex-a-Hop level
can be solved in a given number of steps.
The main idea of the encoding into a SAT CNF formula is to use a set of main
variables indicating the position of the player at each time step, that is to say
the set of three-dimensional vectors $(x, y, t)$ where $x$ is the abscissa of
the position, $y$ the ordinate of the position and $t$ the discrete time step.
The next paragraphs describe the main idea of the encoding by giving the
different type of clauses that must be added to encode the problem. We will
explain the modifications that have to be made in order take in account more
advanced tile types in the following section.

\paragraph{The state clauses}
Obviously, the first set of constraints (clauses) we therefore have to add are
what we've called the {\em state clauses} :
\begin{itemize}
\item At each time step, the player has to be somewhere, on a tile that is
      accessible (no water), therefore we have the clauses containing all the
      possible positions for a given time step t :
      \begin{equation}
      \bigwedge \limits_{t=0}^{t_{max}}
      (\bigvee \limits_{\substack{x \in [\![0, x_{max}]\!]\\
				  y \in [\![0, y_{max}]\!]\\
				  (x, y) \, accessible}}
       (x, y, t))
      \end{equation}

\item But the player can't be at two positions at the same time, therefore we
      have all the clauses containing the negation of two different positions
      for a given time step t :
      \begin{equation}
      \bigwedge \limits_{t=0}^{t_{max}}
      (\bigwedge \limits_{\substack{\{(x, y), (x', y')\} \in
				    \mathcal{P}({[\![0, x_{max}]\!] \times
				    [\![0, y_{max}]\!]})}}
       (\neg(x, y, t) \vee \neg(x', y', t)))
      \end{equation}
\end{itemize}

\paragraph{The movement clauses}
Then, we have to add clauses that are related to the movements of the player.
This means that, for a tile that is accessible (basically isn't water), we have
to get all the neighbor tiles that are accessible from this tile (that is to
say, thanks to one movement). These are essentially the tiles that are at an
infinite-norm distance of one from the considered tile, but there are special
cases to consider when going further in the game, as new type of tiles appear
(this will be developed in the next section). The computation of the accessible
neighbors is done thanks to the function {\em getAccessibleNeighbours} (and the
auxiliary function {\em getNeighbours}). Once we have this function, we define
the {\em movement clauses} :
\begin{equation}
\bigwedge \limits_{t=0}^{t_{max}-1}
       (\bigwedge \limits_{\substack{x \in [\![0, x_{max}]\!]\\
				     y \in [\![0, y_{max}]\!]\\
				     (x, y) \, accessible}}
	((x, y, t) \Rightarrow
	 \bigvee \limits_{(x', y') \, accessible \, from \, (x, y)}
	 (x', y', t+1)))
\end{equation}

\paragraph{The behavioral clauses}
The tiles in a level have different behaviors that influence the movements of
the player, typically (in the basic version) a destroyable field that is
accessed at a given time step mustn't have already been destroyed before. This
gives us the {\em behavioral clauses} :
\begin{equation}
\bigwedge \limits_{\substack{i \in [\![0, x_{max}]\!]\\
			     j \in [\![0, y_{max}]\!]\\
			     (x, y) \, destroyable}}
(\bigwedge \limits_{t=0}^{t_{max}} ((x, y, t) \Rightarrow
				    \neg(x, y, t, d_{destroyed})))
\end{equation}

\paragraph{The start clauses}
Let's then add a clause with a single literal that expresses the fact that the
player must be at the starting point of the level at the time step 0
({\em start clauses}) :
\begin{equation}
(x_{start}, y_{start}, 0)
\end{equation}

\paragraph{The end clauses}
Finally, we must add clauses related to the end of the level, that is to say in
which conditions the level is terminated and was successfully resolved. The
condition is rather simple : there must be no remaining green tile (in terms of
dynamic type, as a tile that isn't initially green, like the turquoise ones, can
become green during the exploration) at the last time step $t_{max}$. Note that
it is not necessary to add a clause for the tiles that aren't destroyable, like
stone or water tiles. This gives us the {\em end clauses} :
\begin{equation}
\bigwedge \limits_{\substack{i \in [\![0, x_{max}]\!]\\
			     j \in [\![0, y_{max}]\!]\\
			     (x, y) \, destroyable}}
\neg(x, y, t_{max}, d_{green})
\end{equation}


\subsubsection{Dealing with advanced tile types}
The two kind of {\em advanced} tiles that we have treated are trampolines, that
appear pretty quickly in the game, and high tiles, that is to say tiles of a
given type (stone, green or turquoise) that are higher than the normal height of
tiles. High tiles of a destroyable type (green or turquoise) have a special
behavior : all the high tiles of a given destroyable type sink down if all the
low tiles of the same type have been destroyed. Naturally, to finish a level,
all the low as well as all the high green tiles must be destroyed.

To take this new type of tiles into account, the first thing that must be
modified is the {\em getAccessibleNeighbours} function :
\begin{itemize}
\item In the case of trampolines, you have to consider all the tiles that are
      accessible by a sequel of jumps on trampolines in a given direction,
      knowing that the final tile on which the player lands must itself be
      accessible (it mustn't obviously be water).
\item In the case of high tiles of type stone that you can't access at all, you
      must precisely not consider them as accessible. Moreover, during the
      computation of the accessible tiles thanks to trampoline jumps, it must be
      taken into account that the player isn't able to jump through these high
      tiles.
\end{itemize}

In our encoding, we consider that when the player does a sequel of jumps on
trampolines, he directly accesses the final tile at the time step at which he
accessed the first trampoline, without considering that the player spent time
"in the air" while jumping (this is justified by the fact that it takes just one
step to go to the first trampoline). This is a problem when it comes to high
stones, as they can perfectly sink down during the time the player jumps from a
trampoline to another tile. To fix that, the fact that a trampoline was accessed
must be remembered thanks to a variable. Therefore we have the {\em trampoline
variables} of the form $trampoline_t$ indicating whether or not a trampoline was
accessed at time step t. The only clauses to add for them are with the
{\em movement clauses} : if coming from the position $(x, y)$ at time step t we
used a trampoline (the infinite-norm distance traveled is greater than 1) to
land on the tile at position $(x', y')$, we have the following implication :
\begin{equation}
(x, y, t) \wedge (x', y', t+1) \Rightarrow trampoline_t
\end{equation}
Else, if we landed on the tile at position $(x', y')$ without using a
trampoline, then we have to add this implication :
\begin{equation}
(x, y, t) \wedge (x', y', t+1) \Rightarrow \neg trampoline_t
\end{equation}

To handle the case of high destroyable tiles, we introduced a new set of
variables indicating if all the low tiles of a given type have been destroyed,
and therefore all the high tiles have sunk down ({\em type height variables}).
In fact, for our case, there are just two of them, $(h_{green}, t)$ and
$(h_{turquoise}, t)$ at a given time step $t$. Therefore, a high destroyable
tile of type $type$ at the position $(x, y)$ can only be accessed at the
time step $t$ if the corresponding variable was already true at the previous
time step if no trampoline was used, or true at the time step $t$ if a trampoline
was used (as the tiles sink down while jumping), which gives us these clauses :
\begin{equation}
((x, y, t) \wedge trampoline_{t-1} \Rightarrow (h_{type}, t))
\wedge
((x, y, t) \wedge \neg trampoline_{t-1} \Rightarrow (h_{type}, t-1))
\end{equation}

Now we must be sure that the {\em type height variables} have the right value by
adding clauses related to the sinking conditions of high tiles. In fact, at a
time step $t$, high tiles of a given type $type$ are sunk down if it already was
the case at the previous time step or if there was no low tile having the same
dynamic type $type$ remaining at the previous time step :
\begin{equation}
\bigwedge \limits_{\substack{i \in [\![0, x_{max}]\!]\\
			     j \in [\![0, y_{max}]\!]\\
			     (x, y) \, not \, high \, of \, type \, type}}
((h_{type}, t) \Rightarrow (h_{type}, t-1) \vee \neg(x, y, t-1, d_{green}))
\end{equation}
The double condition makes sure that tiles which have once sunk down stay sunk down even if there are again low tiles of the respective type (e.g. after visiting a turquoise tile it turns green).


\subsection{How to use the program}

The encoding program was realized in OCaml due to performance reasons. 
The program is called encoder. 
It can be easily build by executing the build.sh script. 
The only requirement is a working OCaml environment. 
To run the program one has to pass three arguments to the executable: path to an input file specifying the map and the starting point, the number of steps for which a solution shall be searched and the path to the output file of the encoding. 
It is important to note that the number of steps defines exactly the number of steps the player has to do in a possible solution. 
That implies that the program will not find a solution which consists of more or less steps. 
The output file will have the DIMACS format. 
Thus, all SAT solvers which support it can be employed. 
The format of the input file will be described in the next section.

The decoding of the result file and the visualization of the solution is done by the python program decoder.py. 
It takes as input parameters the input file describing the map the file with the model generated by the SAT solver and the number of steps of the solution. 
It will then present step by step the path of the solution. 
The only requirement to run the program is a working python environment. 

In order to simplify the execution chain of programs, we also supply the script \emph{hah.sh} which automatically chains all the execution steps necessary to calculate and display the found solution. The input arguments are the input file describing the map and two integers describing the range of steps for which a solution shall be searched. The only thing to adjust is the path to the SAT solver.

\subsubsection{Input file}

The input file starts with a pair of integers indicating the start row and the start column of the player on the map. 
All following lines define the map. 
Each line consists of a sequence of symbols representing the tile types. 
At the moment the following types are supported: \pmb g - Green tile, \pmb s - Stone tile, \pmb t - Turquoise tile, \pmb W - water tile, \pmb j - Trampoline tile. 
For the first three symbols there exist also the capital letter version which denotes that this is a high tile. 
In order to make the map description better readable for humans, all letters have to be separated by a white space. 
Moreover, every odd line has to be shifted by one character to the left. 
That way, one can imitate the hexagonal structure of the field. 
The following example illustrates the explanation:
\begin{verbatim}
2 0
  g
 g g
s G g
 s g
  g	
\end{verbatim}
It is noteworthy that the row and column number starts counting at 0 and that every row and column, even though it might contain a white space, is counted.


\section{Conclusion}

As we have shown in the context of this work, it is possible to solve the games Skyscraper and parts of Hex-a-Hop by reducing them to a SAT problem which then can be efficiently solved by a solver. 
For both problems we have found a polynomial reduction algorithm.
However, we also found out that the encoding of the slightly more complex game Hex-a-Hop produces quickly a huge amount of clauses and variables which can pose a serious challenge for current SAT solvers.
Nevertheless, this is only a question of computing resources and algorithmic techniques and thus only a practical problem and not a fundamental.

\end{document}
